Online {\IterComp}

### 2017-11-02 motivation.tex
changes in motivation

    \paragraph{Using real inputs online is not enough.} We could overcome these problems by just evaluating optimization sequences online,
    while the application is being used. This gives us real data but creates a new problem. Iterative compilation finds good optimizations
    by measuring the efficiency of the resulting binary in terms of speedup over another baseline binary. This requires processing the same
    input in the exact same way twice, one with the optimized and one with the baseline binary, measuring the runtimes, and calculating the
    speedup. When done online, we cannot process the same input twice, each time we process whatever the input happens to be. We need
    another efficiency metric that only requires a single use of an optimized binary.

    \paragraph{Existing efficiency metrics are not up to the task.} A application-specific ones~\cite{alameldeen06,coppa14} may work well,
    but the developer has to choose the appropriate metric and then manually modify the application to calculate it. Existing
    application-agnostic ones do not work. We run each of the \FIXME{32} \textit{KDataSets} benchmarks 1,000 times, each time with a
    different input \textit{and} a different optimized binary. We selected three performance metrics that can be easily measured in a
    single run and can potentially be used as efficiency metrics: runtime, inverse runtime, and
    \textit{Instructions Per Cycle}~(IPC)~\citep{fursin07}.

    \FIXME{Fig.XXX} shows how well these three candidates for an efficiency metric correlate with the metric we would actually like to have,
    speedup over \texttt{-O0}. For all benchmarks, the correlation coefficient~\cite{Bishop:2006:PRM:1162264} is below \FIXME{0.20} for
    runtime and below \FIXME{0.35} for inverse runtime. Both have obvious reasons for being inappropriate efficiency metrics. Doubling the
    amount of data to be processed will cause the runtime to rise and the inverse runtime to fall, even if we use the same binary.
    Efficiency has nothing to do with it. IPC also correlates poorly with performance speedup, the correlation coefficient being just
    \FIXME{0.15}. While it is a common efficiency metric in computer architecture, it works there because it is used with invariant binaries.
    In contrast, any compiler optimization that trades high latency instructions for multiple low latency instructions can reduce IPC while
    improving efficiency.

>>>>

    \paragraph{Using real inputs online is not enough.} We could overcome these problems by just evaluating optimization sequences online
    while the application is being used. This gives us live data but creates a new problem. Optimizations sequences are compared based on
    how much faster the program becomes. When all of them are tested against the same inputs, it is just a matter of comparing runtimes.
    A shorter runtime for processing the same data means a more efficient binary. With live data, we cannot do this, each binary processes
    whatever the input happens to be. We need another efficiency metric that only requires a single use of an optimized binary with a
    single input.

    \paragraph{Existing efficiency metrics are not up to the task.} Application-specific efficiency metrics~\cite{alameldeen06,coppa14} may
    work well, but the developer has to choose the appropriate metric and then manually modify the application to calculate it. Speedup
    over a common baseline binary is a good application-agnostic metric, equating efficiency with how much faster a given input is
    processed. It would be ideal, if we did not have to use the same input twice, one with the baseline and one with the optimized binary.

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose. To show this, we ran
    each of the \FIXME{32} \textit{KDataSets} benchmarks 1,000 times, each time with a different input \textit{and} a different optimized
    binary, and we measured three metrics: runtime, inverse runtime, and IPC. \FIXME{Fig.XXX} shows the correlation coefficient between
    them and a reasonable efficiency metric, speedup over \texttt{-O0}. For all benchmarks, the correlation coefficient is below
    \FIXME{0.20} for runtime and below \FIXME{0.35} for inverse runtime. Both have obvious reasons for being inappropriate efficiency
    metrics. Doubling the amount of data to be processed will cause the runtime to rise and the inverse runtime to fall, even if we use the same binary.
    Efficiency has nothing to do with it. IPC also correlates poorly with efficiency, the correlation coefficient being just \FIXME{0.15}.
    While it is used often in computer architecture, it works there because it is used with invariant binaries. In contrast, any compiler
    optimization that trades high latency instructions for multiple low latency instructions can reduce IPC while improving efficiency.
